---
title: "Huynh_Truc_midterm"
author: "Truc"
date: "3/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(gdata)
library(ggplot2)
library(janitor)   #convert row to column names
library(lubridate) #convert char to date datatype
library(knitr)     #create knitr table
#Set work space
setwd("C:/Users/jacky/Documents/R_Project/Huynh_Truc_midterm")
```


```{r function, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Read file-----------------------------------------------------------
#' 
#' Objective: Function to read the data from the csv file
#'  and store it in local dataframe
#' 
#' Parameter: the file name as a string value(file.name)
#' 
#' Return: un-tidy data frame
#' 
read.data.input <- function(file.name)
{
  un.clean.data <-
    read.csv(
      file.name ,
      header = FALSE,
      quote = "\"",
      stringsAsFactors = FALSE,
      strip.white = TRUE
    )
  
 
  returnValue(un.clean.data)
}

#Create Vector-----------------------------------------------------------
#'
#' Fuction to create sequence vector 
#' 
#' Parmeter: minimum number (min.num)
#' Parmeter: maximum number (max.num)
#' 
#' Return: a vector of column index
vector.creator<-function(min.num, max.num)
{
  cols.nums <- c()
  cols.nums[1] = min.num
  for (i in seq(min.num, max.num, 1)) {
    cols.nums[i] = i + 1
  }
  returnValue(cols.nums)
}
#Clean Data-----------------------------------------------------------
#' 
#' Objective: Get the total number of case in the same country
#' 
#' Parameter: Receive an untidy data frame
#' 
#' Return: return a cleaner dataframe
#' 
get.country.total.case<- function (un.clean.data)
{
  #clear unnecessary column latitude, longtitude, Province, State
  un.clean.data$V3 <- NULL
  un.clean.data$V4 <- NULL
  un.clean.data$V1 <- NULL
  # store the first column to sort the data
  col.data.name <- data.frame(un.clean.data[1,])
  # removing first row to get the sum of each country
  temp.data <- un.clean.data[-(1),]
  #Call vector Creator to create a vector to store column index
  cols.nums <- vector.creator(2, 60)
  #convert char column to integer column
  temp.data[cols.nums] <- sapply(temp.data[cols.nums], as.numeric)
  #sum up all state/province in one country
  clean.data <- aggregate(. ~ V2, temp.data, sum)
  #add the first row back to the datafra
  clean.data <- rbind(col.data.name, clean.data)
  returnValue(clean.data)
}

#Tidy data-----------------------------------------------------------
#' 
#' Objective: tidy the dataframe
#' Objective: adjust all datatype, column names, and values correctly
#' 
#' Parameter: Receive an untidy data frame
#' 
#' Return: return tidy dataframe
#' 
tidy.dataframe <- function(clean.data)
{
  #Change the top row to column
  temp.data <- data.frame(t(clean.data))
  #
  temp.data <-
   data.frame(lapply(temp.data, as.character), stringsAsFactors = FALSE)
  #Change the first value to Date
  temp.data[1, 1] <- 'Date'
  clean.data <- data.frame(temp.data, header = TRUE)
  #Convert first row to column header
  clean.data <- temp.data %>%
    row_to_names(row_number = 1) %>%
    clean_names()
  #Convert first column to date data type
  clean.data <- clean.data %>%
    mutate(date = mdy(date))
  #Create the column vector
  cols.nums <- vector.creator(2, 166)
  # Convert Character into numeric
  clean.data[cols.nums] <- sapply(clean.data[cols.nums], as.numeric)
  returnValue(clean.data)
}

```

<br>
<h2> 1.Introduction</h2>
<h3> 1.1 Summary:</h3>
<p>The data is provided by Johns Hopkins CSSE (CSSEGISandData/COVID-19,n.d.) for the 2019 Novel Coronavirus Visual Dashboard. Also, Supported by ESRI Living Atlas Team and the Johns Hopkins University Applied Physics Lab (JHU APL).</p>

<br>
<h3> 1.2 Variables:</h3>
<p> **Original data frame:** 
  <ul>
    <li>Province/State: Province, State of the following country.</li> 
    <li>Country/Region: Country name.</li> 
    <li>Lat: Latitude of the area.</li> 
    <li>Long: Longtitude of the area.</li> 
    <li>Other column names contain the date of records, column variable contain the number of record by day.</li>
  </ul> 
</p>
<p>
**Tidy data frame:** 
  <ul>
    <li>date: The date of each record.</li>
    <li>Other column contain the number of people of each country's record. </li>
    <li>active: the number of active patient (non death nor recover ) </li>
    <li>recover %: The percentage of recovery patient </li>
    <li>mortality %: The percentage of death patient </li>
    <li>active %: The percentage of COVID patient that active </li>
  </ul>
</p>
 <br>
```{r import.and.tidy}
knitr::opts_chunk$set(echo = FALSE)
#Tidy Confirmed.csv-----------------------------------------------------------
#' 
#' Import and Tidy data from the Confirmed.csv
#'
#Import the data from csv file
un.clean.data <-read.data.input('data/time_series_19-covid-Confirmed.csv')
# Get the sum of case in each country
clean.data <- get.country.total.case(un.clean.data)
# Tidy the data set
confirmed.case <- tidy.dataframe(clean.data)

#Tidy Death.csv-----------------------------------------------------------
#' 
#' Import and Tidy data from the Deaths.csv
#'
#Import the data from csv file
un.clean.data<- read.data.input('data/time_series_19-covid-Deaths.csv')
# Get the sum of case in each country
clean.data <- get.country.total.case(un.clean.data)
# Tidy the data set
deaths.case <- tidy.dataframe(clean.data)

#Tidy Recovered.csv-----------------------------------------------------------
#' 
#' Import and Tidy data from the Recovered.csv
#'
#Import the data from csv file
un.clean.data<- read.data.input('data/time_series_19-covid-Recovered.csv')
# Get the sum of case in each country
clean.data <- get.country.total.case(un.clean.data)
# Tidy the data set
recovered.case <- tidy.dataframe(clean.data)
#Clean up-------------------------------------------------
#'
#' Remove un-use data frame
#' 
rm("un.clean.data", "clean.data")
```

<h2> 2.COVID Worldwide</h2>
<h3> 2.1 Worldwide COVID-19 Trend</h3>
<p>Plot show country around the world affected by COVID since January through March.</p><br>
```{r worldwide.trend, echo=FALSE}
#Create World wide table--------------------------------------------------
#' Calculate worldwide total death, recoved, and confirmed
#'
# Calculate the total death world wide each day
total.death <- data.frame(rowSums(deaths.case[, 2:167]))
# Calculate the total confirmed world wide each day
total.confirmed <- data.frame(rowSums(confirmed.case[, 2:167]))
# Calculate the total recovered case world wide each day
total.recovered <- data.frame(rowSums(recovered.case[, 2:167]))
# Get the date
date.data <- data.frame(confirmed.case[1])
# Combine the 4 data frame to the report day
worldwide.trend <-
  cbind(date.data, total.confirmed, total.death, total.recovered)
colnames(worldwide.trend)<-
  c("date", "confirmed", "death", "recovered") 
#Clean up-------------------------------------------------
#'
#' Remove un-use data frame
#'
rm("total.death", "total.confirmed", "total.recovered","date.data")
#Print plot-------------------------------------------------
#' 2.1
#' Print out plot
#'
ggplot(data = worldwide.trend) +
  geom_line(mapping = aes(x = date,
                            y = confirmed,color="confirmed")) +
  geom_point(mapping = aes(x = date,
                          y = confirmed,color="confirmed"))+
  geom_line(mapping = aes(x = date,
                            y = death,color="death")) +
  geom_point(mapping = aes(x = date,
                          y = death,color="death")) +
  geom_line(mapping = aes(x = date,
                            y = recovered,color="recovered"))+
  geom_point(mapping = aes(x = date,
                          y = recovered,color="recovered"))
```
<br>
<p>Conclusion: while the confirmed case keep increased, the change that the infected person will death is still very low.</p>
<br>
<h3> 2.2 Cases worldwide by date:</h3>
<p> Show first 10 day of the world record since COVID-19 outbreak</p>
```{r worldwide.by.date,echo=FALSE,message=FALSE,warning=FALSE}
#Create worldwide by date---------------------------------------------
#'
#' 2.2 Case world by date
#' Create active column
worldwide.trend$active <- worldwide.trend$confirmed -
  worldwide.trend$death - worldwide.trend$recovered
# Create recovery % column
worldwide.trend$recovery.percent <-  
  round((worldwide.trend$recovered/worldwide.trend$confirmed)*100,3)
# Create mortality % column
worldwide.trend$mortality.percent <-  
  round((worldwide.trend$death/worldwide.trend$confirmed)*100,3)

```

```{r worldwide.trend.table, echo=FALSE,message=FALSE,warning=FALSE}
# Knit table
kable(worldwide.trend,format = "markdown", row.names = FALSE, padding =7, caption = "worldwide case")  
```
<br>
<h3> 2.3 Cases worldwide by country:</h3>
<p> Show infected country order by infected case</p>
```{r world.by.country,message=FALSE,warning=FALSE,echo=FALSE}
#World by country-------------------------------------------------
#'
#' 2.3 Case world by country
#' Create active column
#' #-------------------------------------------------
#'
#' Calculate worldwide total death, recoved, and confirmed
#'
# Calculate the total death world wide each day
total.death <- data.frame(colSums(deaths.case[1:60,-1])) 
# Calculate the total confirmed world wide each day
total.confirmed <- data.frame(colSums(confirmed.case[1:60,-1]))
# Calculate the total recovered case world wide each day
total.recovered <- data.frame(colSums(recovered.case[1:60,-1]))
# Get the country name
df <- tibble::rownames_to_column(total.confirmed, "country")
# add the data together
case.by.country<-cbind(df,total.death,total.recovered)
colnames(case.by.country)<-
  c("country", "confirmed", "death", "recovered") 
#-------------------------------------------------
#' 
#' Create active column
case.by.country$active <- case.by.country$confirmed -
  case.by.country$death - case.by.country$recovered
# Create recovery % column
case.by.country$recovery.percent <-  
  round((case.by.country$recovered/case.by.country$confirmed)*100,3)
# Create mortality % column
case.by.country$mortality.percent <-  
  round((case.by.country$death/case.by.country$confirmed)*100,3)

case.by.country<-case.by.country[ order(-case.by.country$confirmed), ] 

#-------------------------------------------------
#'
#' Remove un-use data frame
#'
rm("total.death", "total.confirmed", "total.recovered","df")

```
```{r case.by.country.table,echo=FALSE}
# Knit table

kable(case.by.country,format = "markdown", row.names = FALSE, padding =7, caption = "worldwide all case by country") 
```
<p> The mortality percent seem lower than the recovery percent in most country. Especially country has the older population. </p>
<br>
<h3> 2.4 Cases worldwide's trend (exclude China):</h3>
<p> Plot to see infected case around the world that exclude China.</p>
```{r world.exclude.china, echo=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------------
#' 
#' 2.4
#' Create new table without china
#' 
#' Calculate worldwide total death, recoved, and confirmed
#'
# exclude china
exclude.china <- function(dataframe)
{
  local.frame<-dataframe
  local.frame$china<-NULL
  # Calculate the total death world wide each day
  local.frame <- 
    data.frame(rowSums(local.frame[, 2:166]))
}

# Calculate the total confirmed world wide each day
total.death<-exclude.china(deaths.case)
total.confirmed <- exclude.china(confirmed.case)
total.recovered <- exclude.china(recovered.case)

# Get the date
date.data <- data.frame(confirmed.case[1])

# Combine the 4 data frame to the report day
worldwide.no.china <-
  cbind(date.data, total.confirmed, total.death, total.recovered)
colnames(worldwide.no.china)<-
  c("date", "confirmed", "death", "recovered") 
#-------------------------------------------------
#'
#' Remove un-use data frame
#'
rm("total.death", "total.confirmed", "total.recovered","date.data")
#-------------------------------------------------
#' 2.4
#' Print out plot
#'
ggplot(data = worldwide.no.china) +
  geom_line(mapping = aes(x = date,
                          y = confirmed,color="confirmed")) +
  geom_point(mapping = aes(x = date,
                           y = confirmed,color="confirmed"))+
  geom_line(mapping = aes(x = date,
                          y = death,color="death")) +
  geom_point(mapping = aes(x = date,
                           y = death,color="death")) +
  geom_line(mapping = aes(x = date,
                          y = recovered,color="recovered"))+
  geom_point(mapping = aes(x = date,
                           y = recovered,color="recovered"))


```
<p>While take out China, the confirmed case is drop dramatically. However the recovered case also drop. Which being said, China has the most infected people, but they are doing pretty good to recover.</p>
<br>
<h3> 2.5 Cases worldwide's total cases exclude China:</h3>
<p> First 10 day of the outbreak in the world that exclude China.</p>

```{r worldwide.no.china, echo=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------------
#'
#' 2.5 Case world by date
#' Create active column
worldwide.no.china$active <- worldwide.no.china$confirmed -
  worldwide.no.china$death - worldwide.no.china$recovered
# Create recovery % column
worldwide.no.china$recovery.percent <-  
  round((worldwide.no.china$recovered/worldwide.no.china$confirmed)*100,3)
# Create mortality % column
worldwide.no.china$mortality.percent <-  
  round((worldwide.no.china$death/worldwide.no.china$confirmed)*100,3)


```

```{r no.china.output, echo=FALSE,message=FALSE,warning=FALSE}
# Knit table
kable(worldwide.no.china,format = "markdown", row.names = FALSE, padding =7, caption = "worldwide case") 
```
<p>Conclusion: The total infected case in China is the highest in the world, but the total deaths in China is not the most (Compare to the 2nd one-Italy)</p>
<br>

<h2> 3.U.S. Analysis</h2>
<h3> 3.1 Trend in the US</h3>
<p> US plot with confired, death, and recovered</p>
```{r us.analys, echo=FALSE,message=FALSE,warning=FALSE}
#---------------------------------------------------------
clean.data.2 <- function(un.clean.data)
{
  newdata <- un.clean.data[ which(un.clean.data$V2=='US'), ]
  temp.data<-data.frame(un.clean.data[1,])
  clean.data<-rbind(temp.data,newdata)
  clean.data$V2<-NULL
  clean.data$V3<-NULL
  clean.data$V4<-NULL
  temp.data <- data.frame(t(clean.data))
  #
  temp.data <-
    data.frame(lapply(temp.data, as.character), stringsAsFactors = FALSE)
  #Change the first value to Date
  temp.data[1, 1] <- 'Date'
  clean.data <- data.frame(temp.data, header = TRUE)
  #Convert first row to column header
  clean.data <- temp.data %>%
    row_to_names(row_number = 1) %>%
    clean_names()
  #Convert first column to date data type
  clean.data <- clean.data %>%
    mutate(date = mdy(date))
  #Create the column vector
  cols.nums <- vector.creator(2, 247)
  # Convert Character into numeric
  clean.data[cols.nums] <- sapply(clean.data[cols.nums], as.numeric)
  returnValue(clean.data)
}

#Import the US confirmed data from csv file
un.clean.data <-
  read.data.input('data/time_series_19-covid-Confirmed.csv')
clean.data <- clean.data.2(un.clean.data)
total.confirmed.us <- clean.data

#Import the US deaths data from csv file
un.clean.data <-
  read.data.input('data/time_series_19-covid-Deaths.csv')
clean.data <- clean.data.2(un.clean.data)
total.deaths.us <- clean.data

#Import the US recovered data from csv file
un.clean.data <-
  read.data.input('data/time_series_19-covid-Recovered.csv')
clean.data <- clean.data.2(un.clean.data)
total.recovered.us <- clean.data

#Create the US summary Cases by date
us.summary.cases <- cbind(
  data.frame(total.recovered.us[, 1]),
  data.frame(rowSums(total.confirmed.us[, 2:247])),
  data.frame(rowSums(total.deaths.us[, 2:247])),
  data.frame(rowSums(total.recovered.us[, 2:247]))
)

# Create the US summary cases by col names
colnames(us.summary.cases)<-
  c("date", "confirmed", "death", "recovered") 

rm("un.clean.data", "clean.data","temp.data")

```

```{r us.plot, echo=FALSE,message=FALSE,warning=FALSE}

#-------------------------------------------------
#' 3.1
#' Print out plot
#'
attach(us.summary.cases)
ggplot(data = us.summary.cases) +
  geom_line(mapping = aes(x = date,
                          y = confirmed,color="confirmed")) +
  geom_point(mapping = aes(x = date,
                           y = confirmed,color="confirmed"))+
  geom_line(mapping = aes(x = date,
                          y = death,color="death")) +
  geom_point(mapping = aes(x = date,
                           y = death,color="death")) +
  geom_line(mapping = aes(x = date,
                          y = recovered,color="recovered"))+
  geom_point(mapping = aes(x = date,
                           y = recovered,color="recovered"))
detach(us.summary.cases)
#-------------------------------------------------
```
<p>Conclusion: while the death case in the us is low, the recover in the us is lower compare to world wide. The infected case is rising faster than most contry</p>

<h3> 3.2 Trend in the US by date</h3>
<p> US record of confired, death, and recovered due to COVID-19 by date</p>
```{r}
#-------------------------------------------------
#'
#' 3.2 Case world by date
#' Create active column
#' 
attach(us.summary.cases)
us.summary.cases$active <- confirmed -death -recovered
# Create recovery % column
us.summary.cases$recovery.percent <-  
  round((recovered/confirmed)*100,3)

# Create mortality % column
us.summary.cases$mortality.percent <-  
  round((death/confirmed)*100,3)
detach(us.summary.cases)


```

```{r}
# Knit table
dt <- us.summary.cases[1:50, ]
kable(dt,format = "markdown", row.names = FALSE, padding =7, caption = "US case by date") 
```
<p>Conclusion: while the death case in the us is low, the recover in the us is much lower compare to world wide. The infected case is rising fast in a very short time (due to weather condition)</p>

<h3> 3.3 Trend in the US by state</h3>
<p> US table by state and province with confired, death, and recovered</p>
```{r, echo=FALSE,message=FALSE,warning=FALSE}
#Create the US summary Cases by state and province
us.summary.cases.by.state <- cbind(
  data.frame(colSums(total.confirmed.us[, -1])),
  data.frame(colSums(total.deaths.us[, -1])),
  data.frame(colSums(total.recovered.us[, -1]))
)

us.summary.cases.by.state <-tibble::rownames_to_column(us.summary.cases.by.state, "state.province")
colnames(us.summary.cases.by.state)<-
  c("state.province", "confirmed", "death", "recovered")
attach(us.summary.cases.by.state )
us.summary.cases.by.state $active <- confirmed -death -recovered
# Create recovery % column
us.summary.cases.by.state$recovery.percent <-  
  round((recovered/confirmed)*100,3)

# Create mortality % column
us.summary.cases.by.state $mortality.percent <-  
  round((death/confirmed)*100,3)
detach(us.summary.cases.by.state)

us.summary.cases.by.state <-us.summary.cases.by.state[ order(-us.summary.cases.by.state$confirmed), ] 
```
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# Knit table
dt <- us.summary.cases.by.state[1:40, ]
kable(dt,format = "markdown", row.names = FALSE, padding =7, caption = "US case by state") 
```


<p>Conclusion: While NewYork has the most confirmed case, Washington state has the most death case. Most big state (New York, Washington, California, New Jersey, Illinois... ) in the US got more infected cases compared to other state</p>

<h2>4. World wide case and the recover change</h2>
```{r}
worldwide.trend$active.percent <-  
  round((worldwide.trend$active/worldwide.trend$confirmed)*100,3)
```

<h3>4.1 World wide Plot</h3>
<p> US table by state and province with confired, death, and recovered</p>
```{r}
ggplot(data = worldwide.trend, aes(x=date, y=active.percent)) +
  geom_bar(stat="identity")
```
<p>Conclusion: The active case was lower down to less that 40% in Marc, but it start raise up again quickly. That mean human body can stay against the virus, but the virus's abilty to spread in the population is also very high</p>

<h3>4.2 Worldwide Record with the active percentage</h3>
<p> Worldwide table by state and province with confired, death, and recovered, active percent</p>
```{r}
kable(worldwide.trend,format = "markdown", row.names = FALSE, padding =7, caption = "US case") 
```
<p>Conclusion: The active case was at 95% when the record start, but it at 66% at the end of the record. At one point, it was lower down to 42%. That mean we can prevent and recover from the virus, but the virus's abilty to spread in the population is also very high</p>

<h2>5. Conclusion</h2>
<h3>5.1 What I learn about COVID-19</h3>
<p>COVID is a worldwide disaster and has kill many people. How ever we can prevent it by keep social distances and personal hygiene.</p>
<h3>5.2 Data vs. Media</h3>
<p>What I learn about COVID-19 is not exactly match the media said. COVID-19 is dangerous but we can still put it under control. (Refer to chart 4.2) At one point we are able to lower the spread wordwide (beginning of March) </p>
<h3>5.3 Advise</h3>
<p>COVID can spread pretty, and it can kill multiple people. These are advice that I learn from the US CDC:<ul>
  <li>Regularly and thoroughly clean your hands with an alcohol-based hand rub or wash them with soap and water.</li>
  <li>Maintain at least 1 metre (3 feet) distance between yourself and anyone who is coughing or sneezing.</li>
  <li>Avoid touching eyes, nose and mouth</li>
  <li>Stay home if you feel unwell. </li>
</ul>
</p>
