---
title: "Introducing to Human Genes and Their Functionality"
author: "Truc"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Objectives
    Explore examples of human genes and their functionality.
### a. What is the CFTR gene?

- The CFTR gene provides instructions for making a protein called the cystic fibrosis transmembrane conductance regulator. The CFTR gene decides the structure of our tissues (bones and muscles), a chemical reaction (breaking down the food we eat), and sends signals around our body (hormones).
- CFTR is a specific gene that codes for a particular protein.

### b. Where that gene is located (which chromosome) and how long is it?

- Located on the long (q) arm of chromosome 7 at position 31.2. The CFTR gene is approximately 189 kb in length and is comprised of 27 exons that encode its genetic sequence (John Hopkins Hospital, n.d.)

### c. What is its function?
- According to John Hopkins Hospital (n.d.), the CFTR gene functions as a "channel across the membrane of cells that produce mucus, sweat, salvia". The channel transports negatively charged particles called "chloride ions into and out of cells". The channel controls the "airways, digestive system, reproductive system, and other organs and tissues." The CFTR protein also regulates the function of other channels, such as those that transport positively charged particles called sodium ions across cell membranes.


### d. Describe a disease-causing mutation in the CFTR gene. Have any gene therapy models been proposed to treat such a disease?
#### 	Disease
According to John Hopkins Hospital (n.d.), one common disease in that CFTR is defective inCystic Fibrosis (CF). This protein is a channel that sits on the surface of cells and transports chloride and other molecules, such as bicarbonate. Mutations in this gene lead to CF. The organs that are typically involved in CF are the skin, pancreas, and lungs.

##### 1.	Sweat Gland
People with CF have very salty sweat. The sweat gland secretes salt and water some of which is typically reabsorbed in the sweat duct. This reabsorption process is markedly abnormal in people with CF. Chloride transport is virtually eliminated because CFTR located on the surface of the cells in the sweat duct is defective. The lack of CFTR function leads to excess chloride in the sweat of people with CF. The high chloride concentration in the sweat can be used to diagnose people with CF.

##### 2.	Lung
The airways are covered with a thin, layer of liquid called airway surface liquid (ASL) and a mucus gel layer. The mucus layer traps bacteria and foreign particles, while cilia on the surface of airway cells constantly move the particles out of the lungs and toward the mouth. This process, called mucociliary clearance is an important defense mechanism that protects the lungs from infection. The ASL also contains antiproteases, antioxidants, antibodies, and other substances that work together to neutralize or destroy invading organisms without damaging the lungs. In CF airways, decreased chloride transport is coupled with excess sodium reabsorption out of the ASL. Since water follows the flow of sodium the ASL and the mucus gel layer become dehydrated.

#### Solution 
When CFTR is defective other channels, including the outwardly rectifying chloride channel (ORCC), the epithelial sodium channel (ENaC), a potassium channel known as ROMK1, and a chloride/bicarbonate exchanger, do not work properly. Also, other chloride channels present on the surface of epithelial cells may be affected in the CF airways. These “alternative” chloride channels have been proposed as a therapeutic target to enhance chloride transport.

### References:
John Hopkins Hospital (n.d.) Coding, construction, and placement of the CFTR protein. Retrieved from: https://hopkinscf.org/knowledge/cftr/#:~:text=Located%20on%20the%20long%20(q,its%20genetic%20sequence%20(1).


## R Overview
- R is an Object Oriented Programming Languages.
In this exercises we try to use R as an calculator.
- User can Find all the math function in R 
```
3*4-5/6
# [1] 11.16667

# Get the directory on the computer
getwd()

# Set the working directory
# setwd("")

print("Hello")

# Math
print (2*pi)
sin (pi/2)

```
```{r seq}
# Multiple argument is done by name then by position
seq(3,20,4)
# Function may define some arguments to have default value
seq(from = 3, to = 20, by = 4)

rnorm(10)
```

## Objects and Variables

- Using an assign operator (<- or =)
- An Object is an abstraction of a memory address used to store value(s).
- Dynamic Programming Languages, happen at runtime. 

## Vector:
- Everything in R is a vector
- Can access individual []
- Access a range of element using (n:m)
```{r vec}
# c is stand for combine in vector
x<-c(2,3,4,5,1)
y=x[2:4]
x<-seq(1,5,2)
y<-x*2
x==y
```
