---
title: "Sequence Statistic for Bio-Informatic"
author: "Truc Huynh"
date: "2/17/2021"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(seqinr)

```
## Objectives
- Explore features of the Bioconductor Packages
- Query/read/Analyze sequence data.

## Description

Write R code to do each of the following tasks:

### Question a

Search the ACNUC “genbank” for the sequences:

#### i. from “Mycobacterium tuberculosis” with accession number JX303316.
``` {r MycobacteriumTuberculosis}
choosebank("genbank")
Q1 <- query("Q1", "AC= JX303316")

```

#### ii. from the “Escherichia coli” with accession number AE005174.
``` {r EscherichiaColi}

Q2 <- query("Q2", "AC= AE005174")

```

### Question b 

Compute the fraction of each base in each sequence.

``` {r ComputeFractionQ1}
# Compute Sequence Q1 (Mycobacterium tuberculosis)

t1 = count(getSequence(Q1$req[[1]]), wordsize = 1)
(t1["a"])/sum(t1)*100
(t1["g"])/sum(t1)*100
(t1["c"])/sum(t1)*100
(t1["t"])/sum(t1)*100

```
``` {r ComputeFractionQ2}

# Compute Sequence Q2 (Escherichia coli)

t2 = count(getSequence(Q2$req[[1]]), wordsize = 1)

(t2["a"])/sum(t2)*100
(t2["g"])/sum(t2)*100
(t2["c"])/sum(t2)*100
(t2["t"])/sum(t2)*100
```

### Question c

For the first sequence, calculate the GC content for each 500-nucleotide
chunks of the sequence. Create a sliding window scattered plot of GC content
using red lines.

``` {r slidingwindowplot}
slidingwindowplot <- function(windowsize, inputseq)
{
  starts <- seq(1, length(inputseq)-windowsize, by = windowsize)
  n <- length(starts) # Find the length of the vector "starts"
  chunkGCs <- numeric(n) # Make a vector of the same length as vector "starts", but just containing zeroes
  for (i in 1:n) {
    chunk <- inputseq[starts[i]:(starts[i]+windowsize-1)]
    chunkGC <- GC(chunk)
    print(chunkGC)
    chunkGCs[i] <- chunkGC
  }
  plot(starts,chunkGCs,type="b",xlab="Nucleotide start position",ylab="GC content")
}
```

``` {r GC_Content}
slidingwindowplot(500, getSequence(Q1$req[[1]]))

```

### Question d

Write a function that finds and plots the count of each codon in a given
sequence. Test your function on both sequences.
``` {r PlotsCodonCounts}
findPotentialCondonAndPlot <- function(sequence)
{
  # Define the codon vector
  CondonTable = count(sequence, wordsize = 3)
  plot(CondonTable,xlab="Codon",ylab="Count")
}

```

Plot of codon's frequency in Sequence 1
``` {r PlotsCodonCountsQ1}
findPotentialCondonAndPlot(getSequence(Q1$req[[1]]))

```

Plot of codon's frequency in Sequence 2
``` {r PlotsCodonCountsQ2}
findPotentialCondonAndPlot(getSequence(Q2$req[[1]]))

```

### Question e

For the second sequence, what are the top three most frequent 3-bp
words?
``` {r 3MostFrequent}
# Get the total count of word size = 3
t2.3 = count(getSequence(Q2$req[[1]]), wordsize = 3)

# Copy of the total count table
t2.3.1 <- t2.3

# Store the index of the most frequent
maxV = {}

#Search for the index of the most frequent
for (i in 1:3)
{
  maxV <- c(maxV,which.max(t2.3.1))
  t2.3.1 <- t2.3.1[-c(which.max(t2.3.1))]
}

# Store the top 3 most frequent
topThreeQ2 = {}

# Store the 3 most frequent 
for (i in maxV)
{
  topThreeQ2 <- c(topThreeQ2,t2.3[i]) 
}

# Display the 3 most frequent
topThreeQ2

# remove all unused data
rm(i, maxV, t2.3, t2.3.1)

```

### Question f 

Write a function to find and return all under-represented DNA words
with a specific length in a given sequence. Test your function using the second
sequence for 2- and 4- bp long words.

#### Function 2 bp long:

``` {r underRepresent}


#################################################
# Function to calculate the 2Bp nucleotides long
#################################################
underRepresent2Bp <- function(sequence)
{
  # This hold the count of 2 character words
  word_frq = count(sequence, 2) / sum(count(sequence, 2))
  
  # This hold the count of 1 character words
  Base_frq = count(sequence, 1) / sum(count(sequence, 1))
  
  # Create a data frame of 2 character words for access sub-string the base
  temp <- data.frame(word_frq)
  
  # container hold the𝜌(Rho) is used to measure
  # how over- or under-represented a particular DNA word is.
  underRepresent.2.Bp = {
  }
  
  for (val in 1:length(word_frq))
  {
    Names <- toString(temp[val, "Var1"]) # Get the nucleotides long names
    
    # Calculate the𝜌(Rho) based on names
    percetage <-
      word_frq[Names] / (Base_frq[substr(Names, 1, 1)] * Base_frq[substr(Names, 2, 2)])
    if (percetage < 1)
    {
      underRepresent.2.Bp <- c(underRepresent.2.Bp, percetage)
    }
  }
  
  View(underRepresent.2.Bp)
  
  return (underRepresent.2.Bp)
}

#################################################
# Function to calculate the 3Bp nucleotides long
#################################################


underRepresent3Bp <- function(sequence)
{
  # This hold the count of 2 character words
  word_frq = count(sequence, 3) / sum(count(sequence, 3))
  
  # This hold the count of 1 character words
  Base_frq = count(sequence, 1) / sum(count(sequence, 1))
  
  # Create a data frame of 2 character words for access sub-string the base
  temp <- data.frame(word_frq)
  
  # container hold the𝜌(Rho) is used to measure
  # how over- or under-represented a particular DNA word is.
  underRepresent.3.Bp = {
  }
  
  for (val in 1:length(word_frq))
  {
    Names <- toString(temp[val, "Var1"]) # Get the nucleotides long
    
    # Calculate the𝜌(Rho) based on names
    percentage <-
      word_frq[Names] / (Base_frq[substr(Names, 1, 1)] * Base_frq[substr(Names, 2, 2)] * Base_frq[substr(Names, 3, 3)])
    if (percentage < 1)
    {
      underRepresent.3.Bp <- c(underRepresent.3.Bp, percentage)
    }
  }
  
  View(underRepresent.3.Bp)
  
  return (underRepresent.3.Bp)
}


#################################################
# Function to calculate the 4Bp nucleotides long
#################################################

underRepresent4Bp <- function(sequence)
{
  # This hold the count of 4 character words
  word_frq = count(sequence, 4) / sum(count(sequence, 4))
  
  # This hold the count of 1 character words
  Base_frq = count(sequence, 1) / sum(count(sequence, 1))
  
  # Create a data frame of 4 character words for access sub-string the base
  temp <- data.frame(word_frq)
  
  # container hold the𝜌(Rho) is used to measure
  # how over- or under-represented a particular DNA word is.
  underRepresent.4.Bp = {
  }
  
  for (val in 1:length(word_frq))
  {
    Names <-
      toString(temp[val, "Var1"]) # Get the four nucleotides long names
    
    # Calculate the𝜌(Rho) based on names by substring each nucleotides
    percentage <-
      word_frq[Names] / (Base_frq[substr(Names, 1, 1)] * Base_frq[substr(Names, 2, 2)]
                         * Base_frq[substr(Names, 3, 3)] * Base_frq[substr(Names, 4, 4)])
    if (percentage < 1) {
      underRepresent.4.Bp <- c(underRepresent.4.Bp, percentage)
    }
  }
  
  View(underRepresent.4.Bp)
  
  return (underRepresent.4.Bp)
}


underRepresent <- function(sequence, SpecLength) {
  if (SpecLength == 4){
    underRepresent4Bp(sequence)    
  } else if (SpecLength == 2){
    underRepresent2Bp(sequence)
  } else if (SpecLength == 3){
    underRepresent3Bp(sequence)
  } else{
    print("Specific Length is undefined")
  }
}

```


``` {r TestUnderRepresent}
# Test the under represent two nucleotides long names
underRepresent(getSequence(Q2$req[[1]]),2)

# test the under present four nucletides long names
underRepresent(getSequence(Q2$req[[1]]),4)

# Test the under represent three nucleotides long names
underRepresent(getSequence(Q2$req[[1]]),3)
```



``` {r close}
closebank()
```

## Notes:

- Handwritten answers are not allowed!

- Use Rmarkdown (https://rmarkdown.rstudio.com/) and provide a neatly
formatted “pdf” file showing both code and output.
• Include your name as a comment at the beginning of the script file.